---
  title: "Classifiers on Table A Classes"
output:
  word_document: default
---

```{r, echo=FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

### Include libraries
```{r, echo=FALSE, warning=FALSE}
library(knitr)
library(e1071)
library(formatR)
library(dplyr)
library(tidyr)
library(ggplot2)    # plots
library(tidyverse)  # data manipulation
library(RSQLite)
suppressMessages(library(dplyr))
suppressMessages(library(pROC))
suppressMessages(library(party))
suppressMessages(library(caret))
suppressMessages(library(xgboost))
suppressMessages(library(randomForest))
suppressMessages(library(miscTools))
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

### Connect to DB
```{r}
conn = dbConnect(RSQLite::SQLite(), "../Data/db.sqlite3")
```

### DB Query Functions
```{r}
a_full <- function() {
  library(RSQLite)
  conn = dbConnect(RSQLite::SQLite(), "../Data/db.sqlite3")
  x = dbGetQuery(conn, "
    SELECT 
      *
    FROM (
      SELECT [Customer.ID], [Customer.Size]
      FROM Customer_Info
    ) AS CI
    JOIN (
      SELECT 
        CustomerLocation, 
        [Date] AS [A.Date],
        [Type] AS [A.Type],
        Year AS [A.Year],
        Month AS [A.Month],
        Day AS [A.Day],
        Occurrences AS [A.Occurrences]
      FROM Table_A_Norm
    ) AS A ON CI.[Customer.ID]=A.CustomerLocation
    JOIN (
      SELECT 
        [Customer.Key], 
        MIN(Requested) AS [B.Min.Requested],
        AVG(Requested) AS [B.Avg.Requested],
        MAX(Requested) AS [B.Max.Requested],
        COUNT(Requested) AS [B.Count.Requested],
        MIN(Early) AS [B.Min.Early],
        AVG(Early) AS [B.Avg.Early],
        MAX(Early) AS [B.Max.Early],
        COUNT(Early) AS [B.Count.Early],
        MIN(Late) AS [B.Min.Late],
        AVG(Late) AS [B.Avg.Late],
        MAX(Late) AS [B.Max.Late],
        COUNT(Late) AS [B.Count.Late],
        MIN([On.Time]) AS [B.Min.On.Time],
        AVG([On.Time]) AS [B.Avg.On.Time],
        MAX([On.Time]) AS [B.Max.On.Time],
        COUNT([On.Time]) AS [B.Count.On.Time],
        MIN(Occurrences) AS [B.Min.Occurrences],
        AVG(Occurrences) AS [B.Avg.Occurrences],
        MAX(Occurrences) AS [B.Max.Occurrences],
        COUNT(Occurrences) AS [B.Count.Occurrences]
      FROM Table_B_Norm
      GROUP BY
        [Customer.Key]
    ) AS B ON A.CustomerLocation=B.[Customer.Key]
    JOIN (
      SELECT 
        Customer, 
        MIN(Type) AS [D.Min.Type],
        AVG(Type) AS [D.Avg.Type],
        MAX(Type) AS [D.Max.Type],
        Count(Type) AS [D.Count.Type],
        MIN(Occurrences) AS [D.Min.Occurrences],
        AVG(Occurrences) AS [D.Avg.Occurrences],
        MAX(Occurrences) AS [D.Max.Occurrences],
        COUNT(Occurrences) AS [D.Count.Occurrences]
      FROM Table_D_Norm
      GROUP BY
        Customer
    ) AS D ON B.[Customer.Key]=D.Customer
    JOIN (
      SELECT 
        Company,
        MIN(Occurrences) AS [E.Min.Occurrences],
        AVG(Occurrences) AS [E.Avg.Occurrences],
        MAX(Occurrences) AS [E.Max.Occurrences],
        COUNT(Occurrences) AS [E.Count.Occurrences]
      FROM Table_E_Norm
      GROUP BY
        Company
    ) AS E ON D.Customer=E.Company
    ORDER BY A.[A.Date] ASC
  ;")
  dbDisconnect(conn)
  drops = c("Customer.ID","CustomerLocation","Customer.Key","Customer","Company","A.Date")
  x = x[, !(names(x) %in% drops)]
  rm(drops)
  
  x$Customer.Size = as.factor(x$Customer.Size)
  x$A.Type = as.factor(x$A.Type)
  
  return(x)
}

a_train <- function() {
  x = a_full()
  index = floor(0.80*nrow(x))
  x = x[1:index,]
  return(x)
}

a_test <- function() {
  x = a_full()
  index = floor(0.80*nrow(x))
  x = x[(index+1):nrow(x),]
  return(x)
}
```

### Split data
```{r}
train = a_train()
test = a_test()
n_classes = length(unique(a_full()$A.Type))
```

### Random Forest predicting Customer.Size
#### Model creation and feature importance
```{r}
set.seed(123) # For reproducibility
rf_model = cforest(A.Type ~ ., data = train, control = cforest_unbiased(mtry=2,ntree=50))
```

#### Get training predictions
```{r}
set.seed(123) # For reproducibility
predicted = predict(rf_model, newdata=train, OOB=TRUE, type="response")
```

#### ROC Curve
```{r}
# ROC curve
rf_cs_auc = auc(as.numeric(train$A.Type), as.numeric(predicted))
rf_cs_auc
```
* AUC for the ROC curve is above.  This implies that the model has that probability that it will be able to distinguish between positive class and negative class.


#### In-sample Error
```{r}
# Compute in-sample results
rf_cs_cm = confusionMatrix(predicted, train$A.Type)
rf_cs_cm
```
* Accuracy is near 99.56%.  This is good except for the fact that it gets Type 1 entirely incorrect.  The dataset is largely disproporionate with 37 type 1's and 8424 type 0's.

#### Get test predictions
```{r}
set.seed(123) # For reproducibility
predicted_test = predict(rf_model, newdata=test, OOB=TRUE, type="response")
```

#### In-sample Error
```{r}
# Compute in-sample results
rf_cs_cm = confusionMatrix(data=predicted_test, reference=test$A.Type)
rf_cs_cm
```
* The test set reveals the same thing as the test set.  The model predicts type 0 very well and completely misclassifies type 1 as type 0.

### Close DB connection
```{r}
dbDisconnect(conn)
```